The Heavy Photon Search Test Detector

### 2013-11-19 hps-testrun-nim.tex
Small rearrangements and edits of ECal perf and trigger perf

The integrated pulse of the each channel was converted to energy by first 
subtracting the pedestal then applying a conversion factor to convert ADC counts to energy. 
The pedestals are measured using special runs where a 100 samples of APD signals, $4$ ns a part, have been recorded at each trigger.  The pedestals were extracted from the 
part of the $400$ ns window before the actual hit in the calorimeter. After subtracting the pedestal, modules with integrated pulse above a threshold are clustered 
using a simple algorithm similar to the one deployed for the trigger, see Sec.~\ref{sec:trigger}. Due to a high readout threshold of $73$~MeV, the average number of modules in a cluster was $\sim 3$ and the simple clustering algorithm worked well for reconstruction of detected shower energy. From pedestal runs the average noise level of the ECal modules were measured to be approximately 15~MeV.

In order to determine conversion factors from ADC counts to energy, the ratios the ECal cluster energy $E$ to the momentum $p$ of  matched tracks in the SVT have been studied in the data and simulation. To compare data and simulation, all inoperable or noisy channels in SVT and ECal have been disabled 
in both data and the simulation. Any efficiency or bias effects that affect the real data should be reflected 
in the simulation. 
%Furthermore, a conversion factor calibration is performed to equalize the gains of  each channel.  simulation to those observed in data by comparing .  The ``weighted $E/p$'' for each crystal is computed, representing the average $E/p$ for clusters that include the crystal
% $\frac{\sum_j w_{j,i}}{\sum_j\frac{P_j}{E_j}w_{j,i}}$.
Iteratively, conversion coefficients have been adjusted until the $E/p$ ratios in the data became similar to the  ones in simulation. The distribution of the $E/p$ ratio in data and simulation can be seen in Fig.~\ref{fig:gains}. The peak position of the distribution shows the sampling fraction of the ECal, the fraction of the incident energy detected in the cluster. The width and tails of the distribution in data indicates imperfect calibration and noise of ECal modules. This level of calibration and the agreement with simulation was found to be sufficient for the multiple scattering analysis using normalized event rates, see Sec.~\ref{sec:mcs}.
\begin{figure}[]
\begin{center}
{\small
	\includegraphics[width=7cm]{figures/eoverp_MC_egs_0_016x0_June5_2013_Sho_Ecal_presentation.png}
	\includegraphics[width=7cm]{figures/eoverp_data_run1351_June5_2013_Sho_Ecal_presentation.png}
	\caption{The ECal energy over track momentum ratio for simulation (top) and data (bottom).\color{red}{\bf (Update style.)}} 
	\label{fig:gains}
}
\end{center}
\end{figure}
%These gains can then be used to convert from the ADC counts reported by the FADC in each channels to the energy deposited in the crystals. In order to find the energy of the incident particle one also needs to know the  the fraction of incident energy that is deposited in the cluster or "sampling fraction". 
%The sampling fraction is parameterized from simulation that show that it is approximately 0.9 for the ECal. The weighted $E/p$ shown in Fig.~\ref{fig:gains} is an approximate measurement of the sampling fraction of that part of the ECal but is position- (somewhat smaller at the edges) and energy-dependent because of the high readout thresholds (explaining the low value compared to 0.9). Since the sampling fraction is not needed to normalize event rates it was not applied in the multiple scattering analysis in Sec.~\ref{sec:mcs}.

\subsubsection{Trigger Performance}
The trigger and DAQ integrate pulses from the ECal modules differently to measure the energy 
deposited in the each crystal. The trigger integrates the pulse using a time-over-threshold window, and 
the DAQ readout integrates using a fixed-size window (5 samples before and 30 samples after a 
threshold crossing). For every event, the trigger reports the trigger decision as a bit mask (top half, 
bottom half or both) and the time the trigger fired. The trigger performance was studied by simulating the 
trigger for each event and comparing to how the events were actually triggered.
First, we simulate the FADC readout board to convert from readout hits (with fixed-size window 
integration) to trigger hits (time-over-threshold integration). Secondly, the CTP clustering 
algorithm and the trigger decision is simulated before we compare the trigger decision and trigger time 
to what was reported by the actual trigger. To eliminate trigger bias, we use a tag and probe method. To 
study the trigger performance in one half of the ECal, we select events which triggered the other half 
and where there was exactly one probe cluster in the ECal half under study. We then measure trigger 
efficiency as the fraction of tagged events that fired the trigger in the probe half as a function of the 
probe cluster energy, shown in Fig.~\ref{fig:turnon}. As a cross-check, the turn-on from the trigger 
threshold was measured to be 1280 in units of ADC counts as expected. The threshold was not 
perfectly sharp because of uncertainties in the conversion from readout to trigger hits described above, 
but based on comparisons with simulation we found that the trigger worked exactly as specified.
\begin{figure}[ht]
\begin{center}
{\small
	\includegraphics[width=7cm]{figures/top_turn_on__bottom_tag_E}
	\caption{Trigger turn-on in the top half of the ECal (bottom is similar) for data and simulation as a 
	function of probe cluster energy.}
\label{fig:turnon}
}
\end{center}
\end{figure}
The trigger turn-on is slow and reaches an uneven plateau just below $1$ GeV for two reasons;  gain 
variations between different crystals lead to the threshold variations and the nonlinearity of the 
time-over-threshold integral means that the effective threshold is higher for clusters that span multiple 
crystals. The effective trigger threshold is therefore dependent on position and energy of the particle 
as well as cluster multiplicity. 
%Overall the trigger appears to have functioned exactly as intended. 
%Changes planned for physics running (constant integration window and per-crystal gain calibration 
%constants for the trigger) will solve both of the issues that led to threshold variations in the Test run.

>>>>

The integrated pulse of the each channel was converted to energy by first 
subtracting the pedestal and then applying a conversion factor to convert ADC counts to energy. 
The pedestals are measured using special runs where each trigger records 100 samples of signals from 
the APDs with 4~ns between each sample. The pedestals were extracted from the 
part of the window before the actual hit in the calorimeter. After subtracting the pedestal, modules with 
integrated pulse above a threshold are clustered using a simple algorithm similar to the one deployed 
for the trigger (see Sec.~\ref{sec:trigger}). Due to the high effective readout threshold of $73$~MeV the 
average number of modules in a cluster was $\sim 3$ and the simple clustering algorithm worked well 
for reconstruction of the detected shower energy. An average noise level of approximately 15~MeV 
was measured in special pedestal runs.

The ratio of the ECal cluster energy $E$ to the momentum $p$ of a matched track in the SVT was used 
to determine the conversion factors from ADC counts to energy. To compare data and simulation, all 
inoperable or noisy channels in the SVT and ECal was disabled in both data and simulation; any 
efficiency or bias that affect the data should be reflected in the simulation. 
%Furthermore, a conversion factor calibration is performed to equalize the gains of  each channel.  simulation to those observed in data by comparing .  The ``weighted $E/p$'' for each crystal is computed, 
representing the average $E/p$ for clusters that include the crystal
% $\frac{\sum_j w_{j,i}}{\sum_j\frac{P_j}{E_j}w_{j,i}}$.
Iteratively, conversion coefficients for each module is adjusted until the $E/p$ ratio in data and 
simulation are similar. The distribution of the $E/p$ ratio in data and simulation can be seen in 
Fig.~\ref{fig:gains}. 
\begin{figure}[]
\begin{center}
{\small
	\includegraphics[width=7cm]{figures/eoverp_MC_egs_0_016x0_June5_2013_Sho_Ecal_presentation.png}
	\includegraphics[width=7cm]{figures/eoverp_data_run1351_June5_2013_Sho_Ecal_presentation.png}
	\caption{The ECal energy over track momentum ratio for simulation (top) and data (bottom).\color{red}{\bf (Update style.)}} 
	\label{fig:gains}
}
\end{center}
\end{figure}
The peak position of the distribution indicates the sampling fraction of the ECal, 
the fraction of the incident particle energy measured in the cluster. The width and tails of the distribution 
in data indicates imperfect calibration and noise of the ECal modules. This level of calibration and the 
agreement with simulation was found to be sufficient for the multiple scattering analysis using 
normalized event rates described in Sec.~\ref{sec:mcs}.

%These gains can then be used to convert from the ADC counts reported by the FADC in each channels to the energy deposited in the crystals. In order to find the energy of the incident particle one also needs to know the  the fraction of incident energy that is deposited in the cluster or "sampling fraction". 
%The sampling fraction is parameterized from simulation that show that it is approximately 0.9 for the ECal. The weighted $E/p$ shown in Fig.~\ref{fig:gains} is an approximate measurement of the sampling fraction of that part of the ECal but is position- (somewhat smaller at the edges) and energy-dependent because of the high readout thresholds (explaining the low value compared to 0.9). Since the sampling fraction is not needed to normalize event rates it was not applied in the multiple scattering analysis in Sec.~\ref{sec:mcs}.

\subsection{Trigger Performance}
The trigger and DAQ integrate pulses from the ECal modules differently to measure the energy 
deposited in the each crystal. The trigger integrates the pulse using a time-over-threshold window, and 
the DAQ readout integrates using a fixed-size window (5 samples before and 30 samples after a 
threshold crossing). For every event, the trigger reports the trigger decision as a bit mask (top half, 
bottom half or both) and the time the trigger fired. The trigger performance was studied by simulating the 
trigger for each event and comparing to how the events were actually triggered.
First, we simulate the FADC readout board to convert from readout hits (with fixed-size window 
integration) to trigger hits (time-over-threshold integration). Secondly, the CTP clustering 
algorithm and the trigger decision is simulated before we compare the trigger decision and trigger time 
to what was reported by the actual trigger. To eliminate trigger bias, we use a tag and probe method. To 
study the trigger performance in one half of the ECal, we select events which triggered the other half 
and where there was exactly one probe cluster in the ECal half under study. We then measure trigger 
efficiency as the fraction of tagged events that fired the trigger in the probe half as a function of the 
probe cluster energy, shown in Fig.~\ref{fig:turnon}. As a cross-check, the turn-on from the trigger 
threshold was measured to be 1280 in units of ADC counts as expected. The threshold was not 
perfectly sharp because of uncertainties in the conversion from readout to trigger hits described above, 
but based on comparisons with simulation we found that the trigger worked exactly as specified.
\begin{figure}[ht]
\begin{center}
{\small
	\includegraphics[width=7cm]{figures/top_turn_on__bottom_tag_E}
	\caption{Trigger turn-on in the top half of the ECal (bottom is similar) for data and simulation as a 
	function of probe cluster energy.}
\label{fig:turnon}
}
\end{center}
\end{figure}
The trigger turn-on is slow and reaches an uneven plateau just below $1$ GeV for two reasons;  gain 
variations between different crystals lead to the threshold variations and the nonlinearity of the 
time-over-threshold integral means that the effective threshold is higher for clusters that span multiple 
crystals. The effective trigger threshold is therefore dependent on position and energy of the particle 
as well as cluster multiplicity. 
%Overall the trigger appears to have functioned exactly as intended. 
%Changes planned for physics running (constant integration window and per-crystal gain calibration 
%constants for the trigger) will solve both of the issues that led to threshold variations in the Test run.
